---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wikimorphemes

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of wikimorphemes is to provide tools for extracting morphemes from Wiktionary entries.
The wikimorphemes api is *not yet stable* and may change.

## Installation

You can install the released version of wikimorphemes from [CRAN](https://CRAN.R-project.org) with:

``` r
# No you can't.
#install.packages("wikimorphemes")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("macmillancontentscience/wikimorphemes")
```

## Examples

The main function in wikimorphemes is `process_word`.

```{r process-word}
library(wikimorphemes)
process_word("unaffable")
process_word("understandable")
process_word("standing")
```

## Contributing to {wikimorphemes}

If you are going to work on this package, please begin by running `download_wikimorphemes_lookup()`. This will download the latest cached version of the lookup (10.7MB), which will then be used in tests.

## Code of Conduct
  
Please note that the wikimorphemes project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). 
By contributing to this project, you agree to abide by its terms.
